Title here

### 2017-01-03 portfolio/text/ms.Rmd
Edits

Individuals who rely on the extraction of natural resources for their
livelihood --- fishers, farmers, miners, and loggers, among others --- face
high levels of income variability driven by a mix of environmental, biological,
and economic variability.
A key component to managing these industries is identifying how regulatory
actions and individual behaviour affect income variability and by extension the
economic stability and sustainable use of natural resources. For commercial
fisheries, recent work has identified that communities and vessels can lower
revenue variability by increasing the diversity of species fished. 
<!-- fix: One sentence clearly stating the general problem being addressed by
this particular study --> 
However, it is unclear if ... extend to individual... year to year changes
<!-- cut down, 1 sentence: -->
Here we identify three axes by which fishers in Alaska, one of the world's most
productive fishing regions, can potentially diversify fishing activities and
investigate how individual actions along each axis affect revenue and revenue
variability. 
<!-- We simultaneously assess the value to individual fishers of choosing to (1) own -->
<!-- permits that allow targeting a diverse fish species assemblage, (2) diversify -->
<!-- by adding additional permits, or (3) catch a greater species diversity within -->
<!-- the constraints of a given permit. --> 
We show that fishing permits with higher species diversity reduces revenue
variability and fishing an additional permit is associated with higher revenue
and lower variability. However, increasing species diversity within the
constraints of a permit set has a fishery-dependent effect on revenue and is
usually (`r g2_dens`% probability) associated with increased revenue uncertainty
the following year. 
Our results demonstrate that the most effective option for individuals to
increase revenue and decrease variability is to purchase permits to participate
in additional or more diverse fisheries. However, this option is costly
and often limited by regulations such as limited entry and catch share
programs.
With increasing environmental variability in a changing climate, it will be
particularly important that individuals who rely on natural resources for
their livelihood have effective strategies to reduce income variability.

@@

Managing natural resources and sustaining the livelihoods of those that depend
on their extraction is hard. These livelihoods tend to have high year-to-year
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability. For example, drought and flooding are a major source of risk for
agricultural food security and farmers' incomes [@gregory2005], and catastrophic
disease outbreaks and wildfires increase risk for the logging industry
[@kirilenko2007].

@@

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio. They can, for
example, target as many species as permitted by a license, own multiple
licenses, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies suggest
there can be substantial benefits to specialization for both farmers and fishers
[@abdulai2000; @arnason2005]. Furthermore, the majority of research examining
the effects of diversification on revenue do so at aggregate levels that combine
individuals (e.g. households, counties, or communities; [@kasperski2013;
@sethi2014]). Thus, key uncertainties remain about the fundamental relationship
between diversity, fishing revenue, and fishing revenue variability at the level
of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. Like much of the rest of the world, the days of open,
unregulated fisheries in most developed countries have passed --- fishers must
own a license to participate in virtually all fisheries [@hilborn2003a]. And
while such regulations have helped reduce overfishing, they limit the available
strategies to reduce income variability. For example, many permits limit fishing
to a single species, diversifying by purchasing additional fishing permits can
be expensive (e.g. ~\$192K USD for a king and tanner crab permit in Southeast
Alaska in 2016; <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some
fisheries limit the entry of new participants through individual fishing quotas
(IFQs) [@grimm2012].

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fished species
diversity [OLE: Call this diversity of catch rather than fished?] on individual
revenue and revenue variability that builds on results
from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability.

@@

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) hold single permits, with a slight increase in single-permit
holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
halibut or salmon permit holders --- compared to other fisheries, halibut permit
holders are more likely to hold halibut plus other permits, whereas salmon
permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
Despite making up only 10--30% of the individuals, multiple permit holders earn
approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean effective
species diversity of this revenue declined somewhat from 1980--2000 but has
since remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely
driven by an increase in salmon specialists in the 1980s and 1990s --- the trend
disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).

@@

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\ref{fig:spoke-all}). For example, individuals fishing a halibut permit plus
one additional permit always had lower expected revenue variability compared to
individuals fishing only a halibut permit (median reduction of `r
100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense (e.g. ~\$192K USD for a king and
tanner crab permit in Southeast Alaska in 2016 REF;
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).

@@

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates with
a major regulatory change. Therefore, we refit our models allowing all
strategy-level effects to vary separately before and after the introduction of
IFQs for strategies involving halibut and sablefish permits. The introduction of
IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals with only
halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

@@

### Implications for limited entry, ITQs, and catch shares

- concerning because in many areas of the world and fisheries, regulations and
  market forces have encouraged specialization / limited (effective)
  diversification (e.g. multiple permits or diverse permits)

- locking into permits limits ability to pick lower risk strategies and is
  getting people locked into low diversity permits that may be doing well now,
  but not necessarily always

@@

\caption{Trends in revenue and diversification over time. (A) Map of fishing
regions in Alaska, USA. (B) Total revenue in billions of 2009 U.S. dollars from
`r range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue)
permit holders. (C) Proportion of single permit holders for all fisheries (thick
black line) and with permit holders omitted each year who fished any of the `r
n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
species fished (Simpson's diversity index) for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the `r n_jacknifed`
top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}

@@

We obtained commercial fisheries landings and revenue data for all permit
holders in Alaska from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan
Commercial Fisheries Entry Commission (CFEC). These data represent a total of
nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r
round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
setting all revenues to 2009 USD [@bea2016]. We implemented a number of
filtering steps to focus on individuals actively engaged in commercial fishing
(Supporting Information). In particular, we removed permit holders whose median
revenue was less than \$10,000 2009 USD to focus on individuals for whom
commercial fishing formed a substantial part of their income.

We used combinations of state and federal fishing permits held by an individual
in in a specific year to define fishing "strategies", representing the species
targeted (Supporting Information). For example, we combined permits that were
otherwise the same except for vessel size (e.g. a longliner fishing sablefish
on a vessel under 60 ft was considered to have the same strategy as a longliner
fishing sablefish on a vessel over 60 ft), and we combined a number of permits
that were targeting a single species and differed only in region fished.
Therefore, for example, someone fishing herring roe in Kodiak and someone
fishing herring roe in Alaska Peninsula were both considered to have a "herring
roe" permit group. For salmon permits, we formed salmon strategies based on
similarities in gear type and species compositon (Supporting Information).

@@

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses [@kasperski2013], we calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
Herfindahl-Hirschman Index) weighted by revenue $R$:

@@

\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, percent change
in species diversity and an indicator variable for increasing or decreasing
species diversity. We exponentiate the equation to ensure all scale parameters,
$\sigma_{i,t}$, are positive. Residuals from the mean component of our model
clearly show a "V"-shaped pattern that the variance component of our model can
capture (Fig. S\ref{fig:resid}). Furthermore, the downside residuals
(year-individual data points with lower than expected revenue) look
qualitatively similar to the complete set of residuals (upside and downside)
(Fig. S\ref{fig:resid-downside}). Therefore, it seemed appropriate to model the
variability as a proxy for revenue risk.

@@

\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
strategy-level predictors. The symbols $M_j$ and $E_j$ represent the mean
species diversity and mean combined fishing season length (effort) for strategy
$j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
strategy-level regression estimating effects across strategies.

We fit our models with Stan 2.14.1 [@standevelopmentteam2016;
@standevelopmentteam2016a] and R 3.3.1 [@rcoreteam2016]. Stan
implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
freedom of 3 and scale of 2) on all $\sigma$ and $\tau$. We ran four chains and
2,000 iterations and discarded the first 1,000 iterations of each chain as warm
up. We checked for chain convergence visually with trace plots, ensured that
$\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
sample size, as calculated by the rstan R package [@standevelopmentteam2016],
was greater than 200 for all parameters [@gelman2014]. We verified that our
model returned sensible estimates by comparing our estimates to an ad-hoc
two-stage model fit with the lme4 R package [@bates2015], where we first fit
a mixed-effects regression model to the mean, and then modelled the absolute
residuals from that first model in a second mixed-effects model representing the
variance component. We also validated the model with simulated data, and were
able to recover the true values (Supporting Information). Code used in our
analysis is available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.

>>>>

Individuals who rely on the extraction of natural resources for their
livelihood --- fishers, farmers, miners, and loggers, among others --- face
high levels of income variability driven by a mix of environmental, biological,
and economic variability.
A key component to managing these industries is identifying how regulatory
actions and individual behaviour affect income variability and by extension the
economic stability and sustainable use of natural resources. For commercial
fisheries, recent work has identified that communities and vessels can lower
revenue variability by increasing the diversity of species fished. 
However, it is unclear if ... extend to individual... year to year changes
Here we identify three axes by which fishers in Alaska, one of the world's most
productive fishing regions, can diversify fishing activities and
investigate how individual actions along each axis affect revenue and revenue
variability. 
We show that fishing permits with higher species diversity reduces individual
revenue variability and fishing an additional permit is associated with higher
revenue and lower variability. However, increasing species diversity within the
constraints of a permit set has a fishery-dependent effect on revenue and is
usually (`r g2_dens`% probability) associated with increased revenue
uncertainty the following year. 
Our results demonstrate that the most effective option for individuals to
increase revenue and decrease variability is to purchase permits to participate
in additional or more diverse fisheries. However, this option is costly
and often limited by regulations such as limited entry and catch share
programs.
With increasing environmental variability in a changing climate, it will be
particularly important that individuals who rely on natural resources for
their livelihood have effective strategies to reduce income variability.

@@

Managing natural resources and sustaining the livelihoods of those that depend
on their extraction is hard. These livelihoods tend to have high year-to-year
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability [@smith1986]. For example, drought and flooding are a major source
of risk for agricultural food security and farmers' incomes [@gregory2005], and
catastrophic disease outbreaks and wildfires increase risk for the logging
industry [@kirilenko2007].

@@

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio. They can, for
example, target as many species as permitted by a license, own multiple
licenses, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies suggest
there can be substantial benefits to specialization for both farmers and fishers
[@smith1986; @abdulai2000; @arnason2005]. Furthermore, the majority of research
examining the effects of diversification on revenue do so at aggregate levels
that combine individuals (e.g. counties or communities;
[@kasperski2013; @sethi2014]). Thus, key uncertainties remain about the
fundamental relationship between diversity, fishing revenue, and fishing
revenue variability at the level of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. The days of open,
unregulated fisheries in most developed countries have passed --- fishers must
own a license to participate in virtually all fisheries [@hilborn2003a]. While
such regulations have helped reduce overfishing, they limit the available
strategies to reduce income variability. For example, many permits limit
fishing to a single species, diversifying by purchasing additional fishing
permits can be expensive (e.g. ~\$192K USD for a king and tanner crab permit in
Southeast Alaska in 2016; <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>),
and some fisheries limit the entry of new participants through individual
fishing quotas (IFQs) [@grimm2012].

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fisheries catch
species diversity on individual revenue and revenue variability that builds on
results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability.

@@

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) hold single permits, with a slight increase in single-permit
holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
halibut or salmon permit holders --- compared to other fisheries, halibut permit
holders are more likely to hold halibut plus other permits, whereas salmon
permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
Despite making up only 10--30% of the individuals, multiple permit holders earn
approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
diversity of this revenue declined somewhat from 1980--2000 but has since
remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely
driven by an increase in salmon specialists in the 1980s and 1990s --- the
trend disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).

@@

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\ref{fig:spoke-all}). For example, individuals fishing a halibut permit plus
one additional permit always had lower expected revenue variability compared to
individuals fishing only a halibut permit (median reduction of `r
100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense (e.g. ~\$192K USD for a king and
tanner crab permit in Southeast Alaska in 2016
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).

@@

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates with
a major regulatory change. Therefore, we refit our models allowing all
strategy-level effects to vary separately before and after the introduction of
IFQs for strategies involving halibut and sablefish permits. The introduction
of IFQs reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with
only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

@@

<!-- ### Implications for limited entry, ITQs, and catch shares -->

<!-- - concerning because in many areas of the world and fisheries, regulations and -->
<!--   market forces have encouraged specialization / limited (effective) -->
<!--   diversification (e.g. multiple permits or diverse permits) -->

<!-- - locking into permits limits ability to pick lower risk strategies and is -->
<!--   getting people locked into low diversity permits that may be doing well now, -->
<!--   but not necessarily always -->

@@

\caption{Trends in revenue and diversification over time. (A) Map of fishing
regions in Alaska, USA. (B) Total revenue in billions of 2009 USD from `r
range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
holders. (C) Proportion of single permit holders for all fisheries (thick black
line) and with permit holders omitted each year who fished any of the `r
n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
species fished (Simpson's diversity index) for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the `r n_jacknifed`
top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}

@@

We obtained commercial fisheries landings and revenue data for all permit
holders in Alaska from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan
Commercial Fisheries Entry Commission (CFEC). These data represent a total of
nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD. We adjusted for
inflation by setting all revenues to 2009 USD [@bea2016]. We implemented
a number of filtering steps to focus on individuals actively engaged in
commercial fishing (Supporting Information). In particular, we removed permit
holders whose median revenue was less than \$10,000 2009 USD to focus on
individuals for whom commercial fishing formed a substantial part of their
income.

We used combinations of state and federal fishing permits held by an individual
in a specific year to define fishing "strategies", representing the species
targeted (Supporting Information). For example, we combined permits that were
otherwise the same except for vessel size (e.g. a longliner fishing sablefish
on a vessel under 60 ft was considered to have the same strategy as a longliner
fishing sablefish on a vessel over 60 ft), and we combined a number of permits
that were targeting a single species and differed only in region fished.
Therefore, for example, someone fishing herring roe in Kodiak and someone
fishing herring roe in Alaska Peninsula were both considered to have a "herring
roe" permit group. For salmon permits, we formed salmon strategies based on
similarities in gear type and species composition (Supporting Information).

@@

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses [@vedeld2007; @illukpitiya2008; @debela2012; @kasperski2013;
sethi2014], we calculated effective species diversity for permit holder $i$ and
time (year) $t$ as the inverse of Simpson's diversity index $\lambda$
[@simpson1949; @jost2006] (also referred to in economics as the
Herfindahl-Hirschman Index) weighted by revenue $R$:

@@

\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, the log ratio
of species diversity from year-to-year and an indicator variable for increasing
or decreasing species diversity. We exponentiate the equation to ensure all
scale parameters, $\sigma_{i,t}$, are positive. Residuals from the mean
component of our model clearly show a "V"-shaped pattern, which the variance
component of our model can capture (Fig. S\ref{fig:resid}). Furthermore, the
downside residuals (year-individual data points with lower than expected
revenue) look qualitatively similar to the complete set of residuals (upside
and downside) (Fig. S\ref{fig:resid-downside}). Therefore, it seemed
appropriate to model the variability as a proxy for revenue risk.

@@

\noindent The $\gamma_{0,j}$ coefficients are modelled with two strategy-level
predictors. The symbols $M_j$ and $E_j$ represent the mean species diversity
and mean combined fishing season length (effort) for strategy $j$. The
coefficients $\eta_1$ and $\eta_2$ represent predictors in this strategy-level
regression estimating effects across strategies.

<!-- TODO: rerun with 2.14.1 -->

We fit our models with Stan 2.14.1 [@standevelopmentteam2016;
@standevelopmentteam2016a] and R 3.3.2 [@rcoreteam2016]. Stan implements the
No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm [@hoffman2014] to
perform Bayesian statistical inference. We assigned weakly informative priors
on all parameters: $N(0, 2^2)$ priors on all $\beta$, $\gamma$, and $\eta$
parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of freedom of 3 and
scale of 2) on all $\sigma$ and $\tau$. We ran four chains and 2,000 iterations
and discarded the first 1,000 iterations of each chain as warm up. We checked
for chain convergence visually with trace plots, ensured that $\hat{R}<1.05$
(the potential scale reduction factor) and that the effective sample size, as
calculated by the rstan R package [@standevelopmentteam2016], was greater than
200 for all parameters [@gelman2014]. We verified that our model returned
sensible estimates by comparing our estimates to an ad-hoc two-stage model fit
with the lme4 R package [@bates2015], where we first fit a mixed-effects
regression model to the mean, and then modelled the absolute residuals from
that first model in a second mixed-effects model representing the variance
component. We also validated the model with simulated data, and were able to
recover the true values (Supporting Information). Code used in our analysis is
available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.
